---
title: "Chapter 2"
author: "Jorge Mendes"
date: "05/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 3

```{r}
set.seed(1014)
library(lobstr)
library(bench)
```

## Quiz 

1 - What are the four common types of atomic vectors? What are the two rare types

```{r}
#Double, character, logical and integer
```

2 - What are attributes? How do you get them and set them?


```{r}
#Additional data in objects.
```

3 - How is a list different from an atomic vector? How is a matrix different from a data frame?

```{r}
#Lists can have multiple types inside it
#All elements in a matrix must be of the same type, in data frame this constraint is only in the columns
```

4 - Can you have a list that is a matrix? Can a data frame have a column that is a matrix?

```{r}
#Yes, adding a dim attribute.
#Yes, just have to make sure that the number of rows is the same of the data frame.
```

5 - How do tibbles behave differently from data frames?

```{r}
#Printing, not turn characters into factors, lazy evaluation and only recycles individual values
```

# Atomic vectors

## Exercises

1 - How do you create raw and complex scalars? (See ?raw and ?complex.)

```{r}
#Raw with as.raw() or charToRaw()
#Complex are created using i, complex() or as.complex()
```

2 - Test your knowledge of the vector coercion rules by predicting the output of the following uses of c():

```{r}
c(1, FALSE)
c("a", 1)
c(TRUE, 1L)

#got the 33 right
```

3 - Why is 1 == "1" true? Why is -1 < FALSE true? Why is "one" < 2 false?

```{r}
1 == "1" #coerced correctly either way

-1 < FALSE #False coerced to numeric (0)

"one" < 2 #2 is converted to string and string comparison is confusing 
```

4 - Why is the default missing value, NA, a logical vector? What’s special about logical vectors? (Hint: think about c(FALSE, NA_character_).)

```{r}
c(FALSE, NA_character_)

?Comparison
#So it is last in the coercion order.
```
5 - Precisely what do is.atomic(), is.numeric(), and is.vector() test for?

```{r}
?is.atomic
#is.atomic tests for the atomic types, not necessarily vectors.

is.atomic(matrix(c(1:10), 2)) #is true

?is.numeric
#more restrictive than is.double and is.integer. Vector needs to be a format that performs common operations. So factor, for example, are excluded.

?is.vector
#is.vector tests if an object has no attribute other than name. 

is.vector(expression(1)) #an expression is considered a vector but a factor, for example, is not because it has other attributes.
```

# Attributes

## Exercises

1 - How is setNames() implemented? How is unname() implemented? Read the source code.

```{r}
setNames

#from the code and documentation it's a shortcut for using names(object) <- names

unname

#turns the names attribute into NULL
```

2 - What does dim() return when applied to a 1-dimensional vector? When might you use NROW() or NCOL()?

```{r}
dim(1:10) #NULL

nrow(1:10) #NULL

ncol(1:10) #NULL

dim
nrow

#nrow and ncol call dim, which calls a primite function
```

3 - How would you describe the following three objects? What makes them different from 1:5?

```{r}
x1 <- array(1:5, c(1, 1, 5))
x2 <- array(1:5, c(1, 5, 1))
x3 <- array(1:5, c(5, 1, 1))

x1
x2
x3

attributes(x1)

#they have a dim attribute that change their behaviour compared to a vector.
```
4 - An early draft used this code to illustrate structure():

```{r}
structure(1:5, comment = "my attribute")
```

But when you print that object you don’t see the comment attribute. Why? Is the attribute missing, or is there something else special about it? (Hint: try using help.)

```{r}
?structure

x_structure <- structure(1:5, comment = "my attribute")

x_structure
attributes(x_structure)

?print.default
?comment

?attributes

#comment attribute is not printed by default.
```

# S3 atomic vectors

## Exercises

1 - What sort of object does table() return? What is its type? What attributes does it have? How does the dimensionality change as you tabulate more variables?

```{r}
?table

#An object of class table. An integer array.

test_table <- table(letters)

typeof(test_table)
attributes(test_table)

#attributes dim, dimnames and class

attr(table(ggplot2::diamonds[c("cut", "color")]), "dim")
attr(table(ggplot2::diamonds[c("cut", "color", "clarity")]), "dim")

#dimensions are the number of unique factors
```

2 - What happens to a factor when you modify its levels?

```{r}
f1 <- factor(letters)
tracemem(f1)
levels(f1) <- rev(levels(f1))

f1
as.numeric(f1)

#the integer associated with the labels change
```

3 - What does this code do? How do f2 and f3 differ from f1?

```{r}
f2 <- rev(factor(letters))
f2
as.integer(f2)

f3 <- factor(letters, levels = rev(letters))
f3
as.integer(f3)

#f2 and f3 have the same integer makeup underneath but the levels have a different order. f1 has the same level order as f3 but with a different integer built.

identical(as.integer(f3), as.integer(f2))
identical(as.integer(f3), as.integer(f1))


identical(levels(f3), levels(f1))
identical(levels(f2), levels(f1))
```

# Lists

## Exercises

1 - List all the ways that a list differs from an atomic vector.

```{r}
#Lists are recursive, elements can be of different types, lists contain references to objects (while vectors have one reference).
```

2 - Why do you need to use unlist() to convert a list to an atomic vector? Why doesn’t as.vector() work?

```{r}
test_list <- list(1:10, 20, 5.3)

test_list

as.vector(test_list)
is.vector(test_list) #already is a vector
#as.vector makes it atomic

?unlist
#unlist gets the values recursively and does the correct testing to get the final vector in a correct format.

test_vector <- unlist(test_list)

typeof(test_vector)
lapply(test_list, typeof)

```

3 - Compare and contrast c() and unlist() when combining a date and date-time into a single vector.

```{r}
today <- Sys.Date()
time_now <- Sys.time()

c(today, time_now)
c(time_now, today)
#c converts for the type of the first element

unlist(list(today, time_now))
unlist(list(time_now, today))
#unlist gets the double underneath the values
```

# Data frames and tibbles

# Exercises

1 - Can you have a data frame with zero rows? What about zero columns?

```{r}
test_df <- data.frame(NULL)

attributes(test_df)

dim(test_df)

#yes
```

2 - What happens if you attempt to set rownames that are not unique?

```{r}
test_df <- data.frame(x = c(1, 2),
                      y = c(3, 4),
                      row.names = c("a", "a"))

#error message
```

3 - If df is a data frame, what can you say about t(df), and t(t(df))? Perform some experiments, making sure to try different column types.

```{r}
test_df <- data.frame(x = c(1, 2),
                      y = c(3, 4))
t(test_df)
class(t(test_df)) #converted to matrix

t(t(test_df))
class(t(t(test_df))) #converted to matrix

test_df <- data.frame(x = c("a", "b"),
                      y = c(3, 4))

t(test_df) #character matrix
class(t(test_df)) #converted to matrix

t(t(test_df)) #character matrix
class(t(t(test_df))) #converted to matrix

?t
#t() and t(t()) return a matrix. Coercion may happen.
```

4 - What does as.matrix() do when applied to a data frame with columns of different types? How does it differ from data.matrix()?

```{r}
?as.matrix
#Values will be coerced according to R coercion hierarchy.

?data.matrix
#data.matrix prioritizes numerical result, coercing dates and factor to numerical. as.matrix prioritize character for these classes, as usual with base R.
```

