---
title: "Chapter 7"
author: "Jorge Mendes"
date: "20/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 7

```{r}
library(rlang)
```

# Quiz

5 - How are <- and <<- different?

```{r}
#<- assigns in the environment it was called, <<- assigns in the global environment.
```

# Environment basics

1- List three ways in which an environment differs from a list.

```{r}
#They're not ordered
#They always have a parent (except empty environment)
#They're not copied when modified
```

2 - Create an environment as illustrated by this picture.

```{r}
loop <- env()
loop$loop <- loop

env_print(loop)
```
3 - Create a pair of environments as illustrated by this picture.

```{r}
loop <- env()
dedoop <- env()

loop$loop <- dedoop
dedoop$dedoop <- loop

env_print(dedoop)
env_print(loop)
```

4 - Explain why e[[1]] and e[c("a", "b")] don’t make sense when e is an environment.

```{r}
#e[[1]] doesn't make sense because environments are not ordered
#e[c("a","b")] doesn't work because [ doesn't work with assignments.
```

5 - Create a version of env_poke() that will only bind new names, never re-bind old names. Some programming languages only do this, and are known as single assignment languages.

```{r}
new_env_poke <- function(env = caller_env(), nm, value, inherit = FALSE, create = !inherit){
  
  if(env_has(env, nm)) stop("Only new names are accepted in this function.")
  
  env_poke(env, nm, value, inherit, create)
}

e3 <- env(x = 1, y = 2)
env_poke(e3, "a", 100)
env_print(e3)

new_env_poke(e3, "a", 200)
new_env_poke(e3, "b", 200)
env_print(e3)

```

6 - What does this function do? How does it differ from <<- and why might you prefer it?

```{r}
rebind <- function(name, value, env = caller_env()) {
  if (identical(env, empty_env())) {
    stop("Can't find `", name, "`", call. = FALSE)
  } else if (env_has(env, name)) {
    env_poke(env, name, value)
  } else {
    rebind(name, value, env_parent(env))
  }
}
rebind("a", 10)
#> Error: Can't find `a`
a <- 5
rebind("a", 10)
a
#> [1] 10

#it change the value bound to a name. It differs from `<<- ` because it only rebinds, while <<- defaults to crating a variable in the global environment if an object is not found. It is preferable because it only does rebinding. 
```

# Recursing over environments

## Exercises

1 - Modify where() to return all environments that contain a binding for name. Carefully think through what type of object the function will need to return.

```{r}
where <- function(name, env = caller_env()) {
  if (identical(env, empty_env())) {
    # Base case
    stop("Can't find ", name, call. = FALSE)
  } else if (env_has(env, name)) {
    # Success case
    env
  } else {
    # Recursive case
    where(name, env_parent(env))
  }
}

where("mean")

e4a <- env(empty_env(), a = 1, b = 2)
e4b <- env(e4a, x = 10, a = 11)

where("a", e4b)


all_where <- function(name, env = caller_env(), env_list = list()) {
  if (identical(env, empty_env())) {
    # Base case
    if(!length(env_list)) stop("Can't find ", name, call. = FALSE)
    env_list
  } else if (env_has(env, name)) {
    # Success case (also recursive)
    env_list <- append(env_list, env)
    all_where(name, env_parent(env), env_list)
  } else {
    # Recursive case
    all_where(name, env_parent(env), env_list)
  }
}

all_where("c", e4b)
all_where("mean")


```

2 - Write a function called fget() that finds only function objects. It should have two arguments, name and env, and should obey the regular scoping rules for functions: if there’s an object with a matching name that’s not a function, look in the parent. For an added challenge, also add an inherits argument which controls whether the function recurses up the parents or only looks in one environment.

```{r}
fget <- function(name, env = caller_env(), inherits = TRUE) {
  if (identical(env, empty_env())) {
    # Base case
    stop("Can't find function ", name, call. = FALSE)
  } else if (env_has(env, name) & is.function(env[[name]])) {
    # Success case
    env
  } else if (inherits){
    # Recursive case
    fget(name, env_parent(env))
  } else {
    stop("Can't find function ", name, call. = FALSE)
  }
}

# debug(fget)
mean <- 3
fget("mean") #expects namespace:base
fget("mean", inherits = FALSE) #expects error since mean is not a function in the caller environment
fget("a", e4b) #expects error since a is not a function

```

# Special environments

## Exercises

1 - How is search_envs() different from env_parents(global_env())?

```{r}
?search_envs
search_envs()

?env_parents
env_parents(global_env())

#search_envs gets the search path, which includes the global environment but not the final empty environment. envs_parents gets all the way above the global environment, excluding it and including the empty environment.
```

2 - Draw a diagram that shows the enclosing environments of this function:

```{r}
f1 <- function(x1) {
  f2 <- function(x2) {
    f3 <- function(x3) {
      x1 + x2 + x3
    }
    f3(3)
  }
  f2(2)
}
f1(1)

#I refuse.
```

3 - Write an enhanced version of str() that provides more information about functions. Show where the function was found and what environment it was defined in.

```{r}
str(median)

improved_str <- function(fn){
  str(fn)
  cat("The function was defined in: ")
  print(fn_env(fn))
  
  cat("The function was found in: ")
  print(fget(as.character(substitute(fn))))
}

improved_str(sum)
```

# Call stacks

## Exercises

1 - Write a function that lists all the variables defined in the environment in which it was called. It should return the same results as ls().

```{r}
local_ls <- function(env=caller_env()){
  env_names(env)
}

local_ls()
ls()

test <- function(){
  a1 <- 1
  b2 <- 2
  local_ls()
}

test()
```

