---
title: "Chapter 9"
author: "Jorge Mendes"
date: "20/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 9

# My first functional: map()

## Exercises

1 - Use as_mapper() to explore how purrr generates anonymous functions for the integer, character, and list helpers. What helper allows you to extract attributes? Read the documentation to find out.


```{r}
library(purrr)

as_mapper(1) #uses pluck
as_mapper("a")
as_mapper(list(1, 2, "a")) 
#all of them uses pluck

View(as_mapper)
methods(as_mapper)

?as_mapper

# attr_getter can be used to get attributes
```

2 - map(1:3, ~ runif(2)) is a useful pattern for generating random numbers, but map(1:3, runif(2)) is not. Why not? Can you explain why it returns the result that it does?

```{r}
map(1:3, ~ runif(2)) #creates an anonymous function that runs for each element in the mapped vector
as_mapper(~runif(2))

map(1:3, runif(2)) #tries to extract elements because the numeric triggers as_mapper numeric method
as_mapper(runif(2))
```

3 - Use the appropriate map() function to:

a. Compute the standard deviation of every column in a numeric data frame.

```{r}
df <- data.frame(x = c(1, 1),
                 y = c(1, 2),
                 z = c(1, 10))

map_dbl(df, sd)
```


b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to do it in two steps.)

```{r}
df <- data.frame(x = c(1, 1),
                 y = c("1", "2"),
                 z = c(1, 10))

map_dbl(df[map_lgl(df, is.numeric)], sd)
```


c. Compute the number of levels for every factor in a data frame.

```{r}
df <- data.frame(x = factor(c("a", "b")),
                 y = factor(c("a", "a")),
                 z = factor(c("b", "b")))

map_dbl(df, ~length(unique(.x)))
```

4 - The following code simulates the performance of a t-test for non-normal data. Extract the p-value from each test, then visualise.

```{r}
trials <- map(1:100, ~ t.test(rpois(10, 10), rpois(7, 10)))

trials_pvalue <- map_dbl(trials, "p.value")

hist(trials_pvalue)
```

5 - The following code uses a map nested inside another map to apply a function to every element of a nested list. Why does it fail, and what do you need to do to make it work?

```{r}
x <- list(
  list(1, c(3, 9)),
  list(c(3, 6), 7, c(4, 7, 6))
)

triple <- function(x) x * 3
map(x, map, .f = triple)
#> Error in .f(.x[[i]], ...): unused argument (function (.x, .f, ...)
#> {
#> .f <- as_mapper(.f, ...)
#> .Call(map_impl, environment(), ".x", ".f", "list")
#> })

#map is passed as an argument to the triple function, which does not make sense.

triple <- function(x) map(x, ~.x * 3)
map(x, triple)
```

6 - Use map() to fit linear models to the mtcars dataset using the formulas stored in this list:

```{r}
formulas <- list(
  mpg ~ disp,
  mpg ~ I(1 / disp),
  mpg ~ disp + wt,
  mpg ~ I(1 / disp) + wt
)

models_mtcars <- map(formulas, ~lm(.x, data = mtcars))
models_mtcars

map(models_mtcars, formula)
```

7 - Fit the model mpg ~ disp to each of the bootstrap replicates of mtcars in the list below, then extract the  R2 of the model fit (Hint: you can compute the R2 with summary().)

```{r}
bootstrap <- function(df) {
  df[sample(nrow(df), replace = TRUE), , drop = FALSE]
}

bootstraps <- map(1:10, ~ bootstrap(mtcars))

get_R2 <- function(x){
  summary(lm(mpg~disp, data = x))$r.squared
}

map_dbl(bootstraps, get_R2)
```

# Map variants

## Exercises

1 - Explain the results of modify(mtcars, 1).

```{r}
modify(mtcars, 1)

a <- modify(mtcars, 1)
identical(a, mtcars)
?modify

str(a)
str(mtcars)
#it will pluck the first element of each column and repeat it, since modify preserves the dimensions.
```

2 - Rewrite the following code to use iwalk() instead of walk2(). What are the advantages and disadvantages?

```{r}
temp <- tempfile()
dir.create(temp)

cyls <- split(mtcars, mtcars$cyl)
paths <- file.path(temp, paste0("cyl-", names(cyls), ".csv"))
walk2(cyls, paths, write.csv)

dir(temp)



temp2 <- tempfile()
dir.create(temp2)

paths <- file.path(temp2, paste0("cyl-", names(cyls), ".csv"))
names(cyls) <- paths
iwalk(cyls, write.csv)

dir(temp2)

#the code is less clear and we lose the group names in the process. But now there's an object clearly named as the file it creates.
```

3 - Explain how the following code transforms a data frame using functions stored in a list.

```{r}
trans <- list(
  disp = function(x) x * 0.0163871,
  am = function(x) factor(x, labels = c("auto", "manual"))
)

nm <- names(trans)
mtcars[nm] <- map2(trans, mtcars[nm], function(f, var) f(var))

#map is run with a function factory that run on the list trans and the selected columns. It creates the function in the list and applies it to a column.
```

Compare and contrast the map2() approach to this map() approach:

```{r}
mtcars[vars] <- map(vars, ~ trans[[.x]](mtcars[[.x]]))

#the map approach does the same, but it iterates over the names of the columns. Which are the same names of the list.
```

4 - What does write.csv() return? i.e. what happens if you use it with map2() instead of walk2()?

```{r}
?write.csv

temp <- tempfile()
dir.create(temp)

cyls <- split(mtcars, mtcars$cyl)
paths <- file.path(temp, paste0("cyl-", names(cyls), ".csv"))
map2(cyls, paths, write.csv) #returns NULL and object names

dir(temp)

```

# Predicate functionals

## Exercises

1 - Why isn’t is.na() a predicate function? What base R function is closest to being a predicate version of is.na()?

```{r}
#is.na returns a value for each element and not just one value. anyNA is the closest to being a predicare version, being any(is.na(.))
```

2 - simple_reduce() has a problem when x is length 0 or length 1. Describe the source of the problem and how you might go about fixing it.

```{r}
simple_reduce <- function(x, f) {
  out <- x[[1]]
  for (i in seq(2, length(x))) {
    out <- f(out, x[[i]])
  }
  out
}

simple_reduce(c(1, 2), sum)
simple_reduce(1, sum)

simple_reduce <- function(x, f) {
  if(length(x) <= 1) return(x)
  
  out <- x[[1]]
  for (i in seq(2, length(x))) {
    out <- f(out, x[[i]])
  }
  out
}

simple_reduce(c(1, 2), sum)
simple_reduce(1, sum)

```

3 - Implement the span() function from Haskell: given a list x and a predicate function f, span(x, f) returns the location of the longest sequential run of elements where the predicate is true. (Hint: you might find rle() helpful.)

```{r}
span <- function(x, f){
  rle_result <- map(keep(x, f), rle)
  rle_result
}


span(list(c(1, 2, 3), c(4, 4, 5), c(3, 2, 2 ,2)), function(x) all(x > 1))

##seee in results since I don't understand what span does
```

4 - Implement arg_max(). It should take a function and a vector of inputs, and return the elements of the input where the function returns the highest value. For example, arg_max(-10:5, function(x) x ^ 2) should return -10. arg_max(-5:5, function(x) x ^ 2) should return c(-5, 5). Also implement the matching arg_min() function.

```{r}
arg_max <- function(x, f){
  mapped_values <- map_dbl(x, f)
  x[mapped_values == max(mapped_values)]
}

arg_max(-10:5, function(x) x ^ 2)
arg_max(-5:5, function(x) x ^ 2)

arg_min <- function(x, f){
  mapped_values <- map_dbl(x, f)
  x[mapped_values == min(mapped_values)]
}

arg_min(-10:5, function(x) x ^ 2)
arg_min(-5:5, function(x) x ^ 2)
```

5 - The function below scales a vector so it falls in the range [0, 1]. How would you apply it to every column of a data frame? How would you apply it to every numeric column in a data frame?

```{r}
scale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}

apply(mtcars, 2, scale01) #turns into matrix
map_dfc(mtcars, scale01) #rownames are lost
modify(mtcars, scale01) #keeps it as a data.frame

identical(as.data.frame(apply(mtcars, 2, scale01)), modify(mtcars, scale01)) #TRUE

apply(Filter(is.numeric, iris), 2, scale01) #drops numeric variables
map_if(iris, is.numeric, scale01) #lists
modify_if(iris, is.numeric, scale01) #great
```

# Base functionals

## Exercises

1 - How does apply() arrange the output? Read the documentation and perform some experiments.

```{r}
?apply
#not the answers but it coerces te object to a matrix

#if each call returns a vector with length greater than one apply will return an array. If n is one it will return a vector or an matrix. If values don't have the same length it returns a list.

#results are coerced to one of the basic vector types. Factor, for example, are coerced to character.
```

2 - What do eapply() and rapply() do? Does purrr have equivalents?

```{r}
?eapply
#apply function over values in an environment and returns them in a list. I think imap could lead to the same result.

?rapply
#recursive and more flexible version of apply. I think it's modify.
```

