---
title: "Chapter 5"
author: "Jorge Mendes"
date: "20/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 5

```{r}
set.seed(1014)
```

## Quiz 

1 - What is the difference between if and ifelse()?

```{r}
#Ifelse is vectorized, which may cause recycling issued. If expects one element, if given more it uses the first.
```

2 - In the following code, what will the value of y be if x is TRUE? What if x is FALSE? What if x is NA?

```{r}
x <- NA
y <- if (x) 3
y

#3, NULL, ERROR.
```
3 - What does switch("x", x = , y = 2, z = 3) return?

```{r}
# 2

switch("x", x = , y = 2, z = 3)
```

# Choices

## Exercises

1 - What type of vector does each of the following calls to ifelse() return?

Read the documentation and write down the rules in your own words.

```{r}
ifelse(TRUE, 1, "no") #numeric
ifelse(FALSE, 1, "no") #character
ifelse(NA, 1, "no") # NA (logical)

typeof(ifelse(NA, 1, "no"))

#ifelse returns an element of the same length as test and the type of yes or no, or a missing value, if a missing value is provided. If answer has multiple types they will be coerced, preferably to the type of the yes argument.
```

2 - Why does the following code work?

```{r}
x <- 1:10
if (length(x)) "not empty" else "empty"
#> [1] "not empty"

x <- numeric()
if (length(x)) "not empty" else "empty"
#> [1] "empty"

#zero is coerced to FALSE and a non zero number is coerced to TRUE.
```

# Loops

## Exercises

1 - Why does this code succeed without errors or warnings?

```{r}
x <- numeric()
out <- vector("list", length(x))
for (i in 1:length(x)) {
  out[i] <- x[i] ^ 2
}
out

# 1:length(x) results in 1 0 because `:` also works backwards. This makes the loops run for the first element of x, which doesn't exist and results in an NA in out.
```

2 - When the following code is evaluated, what can you say about the vector being iterated?

```{r}
xs <- c(1, 2, 3)
tracemem(xs)
for (x in xs) {
  xs <- c(xs, x * 2)
  print(tracemem(xs))
}
xs
print(tracemem(xs))
#> [1] 1 2 3 2 4 6

#the vector is doubled and iteration is happening on a copy of xs made when entered on the loop.
```


3 - What does the following code tell you about when the index is updated?

```{r}
for (i in 1:3) {
  i <- i * 2
  print(i) 
}
#> [1] 2
#> [1] 4
#> [1] 6

#index is assigned in the beggining of each loop.
```

