---
title: "Chapter 4"
author: "Jorge Mendes"
date: "20/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 4

```{r}
set.seed(1014)
```

## Quiz 

1 - What is the result of subsetting a vector with positive integers, negative integers, a logical vector, or a character vector?

```{r}
#Positive integers - Gets the values in the supplied position
#Negative integers - Removes the values in the supplied position
#Logical vector - Return the values when TRUE
#Character vector - Return values whose names matches the character
```

2 - What’s the difference between [, [[, and $ when applied to a list?

```{r}
# [ always returns a list. Can be used for multiple elements.
# [[ returns a single element in the defined position.
# $ is a shorthand for [[.
```

3 - When should you use drop = FALSE?

```{r}
#When you don't want the object to be simplified. Useful inside functions.
```

4 - If x is a matrix, what does x[] <- 0 do? How is it different from x <- 0?

```{r}
# x[] <- 0 keeps dimensions and changes all elements to 0
#x <- 0 bounds the name x to the value 0
```

5 - How can you use a named vector to relabel categorical variables?

```{r}
# Using the name vector as for subsetting could be used for this.
```

# Selecting multiple elements

## Exercises

1 - Fix each of the following common data frame subsetting errors:

```{r}
mtcars[mtcars$cyl = 4, ]
mtcars[mtcars$cyl == 4, ]

mtcars[-1:4, ]
mtcars[-c(1:4), ] #or
mtcars[1:4, ]

mtcars[mtcars$cyl <= 5]
mtcars[mtcars$cyl <= 5, ]


mtcars[mtcars$cyl == 4 | 6, ]
mtcars[mtcars$cyl %in% c(4, 6), ] #or
mtcars[mtcars$cyl == 4 | mtcars$cyl == 6, ]
```

2 - Why does the following code yield five missing values? (Hint: why is it different from x[NA_real_]?)

```{r}
x <- 1:5
x[NA]

x[NA_integer_]

#NA is a logical, so it's recicled to the length of the vector.
```

3 - What does upper.tri() return? How does subsetting a matrix with it work? Do we need any additional subsetting rules to describe its behaviour?

```{r}
x <- outer(1:5, 1:5, FUN = "*")
x[upper.tri(x)]

?upper.tri
#upper.tri returns a logical matrix with TRUEs in the upper triangle of a matrix

#subsetting a matrix with it drops dimension and returns the TRUE values as vectors
#Apparently the logical matrix works as a vector subsetting and not as a matrix. In `[`help section and in the book the matrix subsetting only refers to integer or character matrices. Maybe a new rule is needed that says that logical matrices are coerced to vectors.
```

4 - Why does mtcars[1:20] return an error? How does it differ from the similar mtcars[1:20, ]?

```{r}
mtcars[1:20]
#looks for columns. mtcars only has 11 columns, thus the error when selecting 20

mtcars[1:20, ]
#now the subsetting happens on rows, where there are more than 20 to select
```

5 - Implement your own function that extracts the diagonal entries from a matrix (it should behave like diag(x) where x is a matrix).

```{r}
x <- outer(1:5, 1:4, FUN = "*")
x
diag(x)

my_diag <- function(x){
  apply_dimension <- min(dim(x))
  result_diag <- vector(mode = typeof(x), length = apply_dimension)
  for(i in seq_len(apply_dimension)){
    result_diag[i] <- x[i, i]
  }
  result_diag
}

my_diag(x)
```

6 - What does df[is.na(df)] <- 0 do? How does it work?

```{r}
df <- data.frame(x = c(NA,2),
                 y = c(1, NA))
df[is.na(df)] <- 0

df
#change NA values to 0

df <- data.frame(x = c(NA,2),
                 y = c(1, NA))
is.na(df)
#changes where position is TRUE for the logical is.na test. This confirms that logical subsetting works as vectors.
```

# Selecting a single element

## Exercises

1 - Brainstorm as many ways as possible to extract the third value from the cyl variable in the mtcars dataset.

```{r}
mtcars$cyl[3]
mtcars[3, "cyl"]
mtcars[["cyl"]][3]
mtcars[3, ]$cyl
mtcars[3, ][["cyl"]]
```

2 - Given a linear model, e.g., mod <- lm(mpg ~ wt, data = mtcars), extract the residual degrees of freedom. Then extract the R squared from the model summary (summary(mod))

```{r}
mod <- lm(mpg ~ wt, data = mtcars)

mod
mod$df.residual

attributes(summary(mod))
summary(mod)$r.squared
```

# Subsetting and assignment

## Exercises

1 - How would you randomly permute the columns of a data frame? (This is an important technique in random forests.) Can you simultaneously permute the rows and columns in one step?

```{r}
df <- data.frame(x = c(1:5),
                 y = c(6:10),
                 z = c(11:15))
df

df[sample(NCOL(df))]
df[sample(NROW(df)), sample(NCOL(df))]
```

2 - How would you select a random sample of m rows from a data frame? What if the sample had to be contiguous (i.e., with an initial row, a final row, and every row in between)?

```{r}
df <- data.frame(x = c(1:5),
                 y = c(6:10),
                 z = c(11:15))

df[sample(NROW(df), 2), ] #m = 2

initial <- 2
end <- 5
df[c(initial, sample((initial+1):(end-1), 1), end), ]
```

3 - How could you put the columns in a data frame in alphabetical order?

```{r}
df <- data.frame(c = c(1:5),
                 a = c(6:10),
                 b = c(11:15))
df[sort(names(df))]
```

