---
title: "Chapter 18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rlang)
library(lobstr)
```

# Abstract syntax trees

## Exercises

1 - Reconstruct the code represented by the trees below:

#> 
[33m█
[39m─
[1m
[35mf
[39m
[22m 
#> └─
[33m█
[39m─
[1m
[35mg
[39m
[22m 
#>   └─
[33m█
[39m─
[1m
[35mh
[39m
[22m
#> 
[33m█
[39m─
[1m
[35m`+`
[39m
[22m 
#> ├─
[33m█
[39m─
[1m
[35m`+`
[39m
[22m 
#> │ ├─1 
#> │ └─2 
#> └─3
#> 
[33m█
[39m─
[1m
[35m`*`
[39m
[22m 
#> ├─
[33m█
[39m─
[1m
[35m`(`
[39m
[22m 
#> │ └─
[33m█
[39m─
[1m
[35m`+`
[39m
[22m 
#> │   ├─
[1m
[35mx
[39m
[22m 
#> │   └─
[1m
[35my
[39m
[22m 
#> └─
[1m
[35mz
[39m
[22m

```{r}
#I think there are some missing characters so I'll skip it
```

3 - What’s happening with the ASTs below? (Hint: carefully read ?"^".)

```{r}
lobstr::ast(`x` + `y`)
#> █─`+` 
#> ├─x 
#> └─y
lobstr::ast(x ** y)
#> █─`^` 
#> ├─x 
#> └─y
lobstr::ast(1 -> x)
#> █─`<-` 
#> ├─x 
#> └─1

# Values are translated by the parser
```

4 - What is special about the AST below? (Hint: re-read Section 6.2.1.)

```{r}
lobstr::ast(function(x = 1, y = 2) {})
#> █─`function` 
#> ├─█─x = 1 
#> │ └─y = 2 
#> ├─█─`{` 
#> └─<inline srcref>

#It separates the formals from the body
```

5 - What does the call tree of an if statement with multiple else if conditions look like? Why?

```{r}
lobstr::ast(
  if(x ==1 ){
    1
  } else if(x == 2){
    2
  } else {
    if(x < 0) -1
    2
  }
)

#Very long and nested tree. And apparently else is not a function (or at least is not called.)
```

# Expressions

## Exercises

1 - Which two of the six types of atomic vector can’t appear in an expression? Why? Similarly, why can’t you create an expression that contains an atomic vector of length greater than one?

```{r}
?rlang::expr

typeof(rlang::expr(1L))
typeof(rlang::expr(1i))
typeof(rlang::expr(charToRaw("a"))) #call

#Raw and have no idea about the other

typeof(rlang::expr(c(1, 2, 3)))
rlang::is_call(rlang::expr(c(1, 2, 3)))
#Because you have to use c, turning it into a call
```
2 - What happens when you subset a call object to remove the first element? e.g. expr(read.csv("foo.csv", header = TRUE))[-1]. Why?

```{r}
rlang::expr(read.csv("foo.csv", header = TRUE))[-1] 
#it creates another call object in which foo is the function name
```
3 - Describe the differences between the following call objects.

```{r}
x <- 1:10

call2(median, x, na.rm = TRUE) #median and 1:10 are called
call2(expr(median), x, na.rm = TRUE) #median is not called, x is
call2(median, expr(x), na.rm = TRUE) #median is called, x is not
call2(expr(median), expr(x), na.rm = TRUE) #median and x are not called
```

4 - rlang::call_standardise() doesn’t work so well for the following calls. Why? What makes mean() special?

```{r}
call_standardise(quote(mean(1:10, na.rm = TRUE)))
#> mean(x = 1:10, na.rm = TRUE)
call_standardise(quote(mean(n = T, 1:10)))
#> mean(x = 1:10, n = T)
call_standardise(quote(mean(x = 1:10, , TRUE)))
#> mean(x = 1:10, , TRUE)

#mean has only one argument: x, the other one is ..., which doesn't work well with call_standardise
```

5 - Why does this code not make sense?

```{r}
x <- expr(foo(x = 1))
x

names(x) <- c("x", "y")

x$y

#you can't name the values but it is not so useful. And x in the function will be named y.
```

6 - Construct the expression if(x > 1) "a" else "b" using multiple calls to call2(). How does the code structure reflect the structure of the AST?

```{r}
call2("if", call2(">", x, 1), "a", "b")
lobstr::ast(if(x > 1) a else b)
#the call represents well the ast structure
```
# Parsing and grammar

## Exercises

1 - R uses parentheses in two slightly different ways as illustrated by these two calls:

```{r}
lobstr::ast(f((1))) #parenthesis is a function that holds the value
lobstr::ast(`(`(1 + 1)) #parentheses is a function that makes sure the function inside is executed
```
Compare and contrast the two uses by referencing the AST.

2 - = can also be used in two ways. Construct a simple example that shows both uses.

```{r}
lobstr::ast(x = 3 == 5)
```

